### 2022-04-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@038b5ced3967567216335ab65613b6abe255f450 🚀

\vfill
Coming soon

@@

Despite the difference in evolutionary distinctiveness globally, there
are reasons to expect that the phylogeography of bats and
betacoronaviruses should show some degree of congruence. High density of
hosts sharing the same virus (albeit possibly different strains) can
drive or result from evolution of the bat antiviral immune system,
resulting in spatially distinct immunological responses, as evidenced in
several bat species (Banerjee et al. 2020). Immune characteristics that
allow bats to be better adapted to infection by emerging viruses
(Gorbunova et al., 2020; Irving et al., 2021) may be related to a wide
variety of diets (Jones et al., 2022; Moreno Santillán et al., 2021;
Banerjee et al., 2020; Schneeberger et al., 2013), themselves likely to
be driven by spatial effect, especially at the local scale -- bats,
indeed, occupy a variety of environments.

@@

\begin{figure}
\hypertarget{fig:risk}{%
\centering
\includegraphics{figures/risk_map.png}
\caption{This is the legend of the figure\ldots{}}\label{fig:risk}
}
\end{figure}

@@

Based on the previous result, we extracted the yellow component of the
risk map (TP add methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in Fig. xxA. However, this
maps the potential risk, which must be weighed by the potential for
contacts with humans. As a proxy for this measure, we used the
proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et
al., 2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in Fig. xxB.
Most of South America and Europe are at low risk, as although densely
populated, settlements tend to be in areas with lower potential risk.
However, this mapping reveals that South-East Asia, the Indian
subcontinent, and parts of sub-Saharan Africa, are at high risk due to
the overlap between built areas and bat communities representing more
opportunities for cross-species transmission of betacoronaviruses.

\begin{figure}
\hypertarget{fig:compound}{%
\centering
\includegraphics{figures/risk_compounded.png}
\caption{This is the legend of the figure\ldots{}}\label{fig:compound}
}
\end{figure}

@@

Our study focuses largely on the biogeography of hosts. Yet, we know
that viruses with high host plasticity, that is, the ability of a given
virus to adapt to various taxonomic orders and ecological groups
(Kreuder Johnson et al., 2015); are more likely to amplify viral
spillover, followed by secondary human-to-human transmission, and
geographical spread (Hazarie et al., 2021). High viral host plasticity
is an especially important trait for RNA viruses such as betacov
(Kreuder Johnson et al., 2015; Haddad et al., 2021). Indeed, our
analysis of viral sequences reveals that Latin America is a hotspot of
viral distinctiveness, suggesting that this part of the bats-betacov
system may be undergoing independent evolutionary dynamics (related
species sharing viruses that are different from the rest of the global
pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
richness is high but sharing is low; this suggests a different type of
evolutionary dynamics (unrelated viruses coevolving with evolutionarily
distinct hosts, generating high diversity locally).

Driven by the need to understand the ecological factors involved in the
emergence of viral pathogens, we spatially mapped bat-betacoronavirus
interactions worldwide, using (i) a database of known betacov
hosts(Becker et al., 2020), and (ii) range maps for the hosts according
to IUCN (IUCN 2021). To reflect the fact that the risk posed by viruses
has many ecological origins, we quantified the phylogenetic diversity of
hosts, their compositional uniqueness, and the expected viral sharing.
Because these components of risk matter when contrasted to human
density, we compared them to a proxy, namely the proportion of each
pixel that is covered by urban or built land. This provides a synthetic
risk map, allowing to identifying of hotspots where the
bat-betacoronavirus system may originate viruses in humans. SE Asia is
one of the regions with the highest risk since, according to our
results, several of its conditions could increase the risk of
transmission of the virus.

Species richness, therefore, is not a sufficient measure of viral risk.
This is exemplified in our results, where both South America and
South-Eastern Asia have a high species richness of betacov hosts, but
only the latter region has a high risk. Specifically, because previous
studies propose that Asia is important when it comes to understanding
the evolutionary origin of various mammalian taxa (Beard C K, 1988).

There are several factors that drive changes in the diversity of bats
(Alves et al., 2018), but human activities' effects on the ecosystem
(like modifications of land use) could significantly decrease it.
Therefore, it can be suggested that changes in the diversity of betacovs
in bats are linked to their biogeographic variation, and human
population density and other anthropogenic factors are decisive
moderators for its implications in public health. With the increase of
contact between humans and potential hosts, we also increase the risk of
emergence of novel diseases (Johnson et al., 2020), as previous studies
on RNA viruses suggest the importance of host phylogeography at the time
of virus dispersal (Gryseels et al., 2017).

This diversity of hosts and how the exchange of viruses occurs between
species, is largely affected by the different environmental changes, as
the case of sarbecovirus bats reservoirs (Muylaert et al., 2021) where
they are affected by the area of the cave or the alteration of the
forest, which could result in modifications of host distribution.
Additionally, our results highlight the importance of Asia as a betacov
hotspot, which is consistent with recent studies (Muylaert et al.,
2021), where projections on this area suggest that new future events of
sarbecovirus viral exchange might be easily spread among species or
humans.

One of these scenarios where interaction between bats and humans can
occur can be seed dispersal in tropical agroecosystems. It opens the
discussion of whether the fruits thrown by bats not only disperse seeds
but could also be a source of indirect interaction between viruses of
bat origin and humans (Deshpande et al., 2022) . This represents a
challenge for conservation strategies and disease ecology since we have
areas with potential zoonotic viruses and bat-human interaction.
However, it must still be taken into account the quantification of real
exposure from several scenarios, where there can be directly or
indirectly bat - human interaction.

This probability involves multiple factors, among which the relatedness
of hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
of hosts within a locality to share viruses, which may limit viral
diversity because of within-host competition (Leeks et al., 2018;
Sallinen et al., 2020). All things considered, the richness of known
betacoronaviruses hosts is not a sufficient predictor of spillover risk,
and we should move towards a more coevolutionary understanding of this
system.

Considering whether viruses easily adapted to multiple hosts have lower
virulence on these hosts, or lower ability to jump to hosts with
different immune characteristics, should yield valuable additional
predictors for the total risk of spillover.

>>>>

\vfill
Driven by the need to understand the ecological factors involved in the
emergence of betacoronavirus through bat hosts, we develop an approach
to the assesment of spillover risk based on the Geographic Mosaic Theory
of Coevolution. In doing so, we provide a global mapping of the risk
posed by viruses, reflecting the fact that this risk has many ecological
and evolutionary origins. Our framework reveals that reservoir richness
alone is not a sufficient predictor of risk, and offers alternative
insights based on viral sharing, host compositional uniqueness, and host
phylogenetic diversity. By comparing our aggregated measure of risk to a
proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
system may originate novel viruses in humans.

@@

Despite the difference in evolutionary distinctiveness globally, there
are reasons to expect that the phylogeography of bats and
betacoronaviruses should show some degree of congruence. High density of
hosts sharing the same virus (albeit possibly different strains) can
drive or result from evolution of the bat antiviral immune system,
resulting in spatially distinct immunological responses, as evidenced in
several bat species (Banerjee et al. 2020). Immune characteristics that
allow bats to be better adapted to infection by emerging viruses
(Gorbunova et al., 2020; Irving et al., 2021) may be related to a wide
variety of diets (Jones et al., 2022; Moreno Santillán et al., 2021;
Banerjee et al., 2020; Schneeberger et al., 2013; Muylaert et al.,
2021), themselves likely to be driven by spatial effect, especially at
the local scale -- bats, indeed, occupy a variety of environments.

@@

\begin{figure}
\hypertarget{fig:risk}{%
\centering
\includegraphics{figures/risk_map.png}
\caption{Extraction of a measure of risk based on the colorimetric space
from fig.~\ref{fig:trivariate}. The risk is a composite measure of the
color value and angular distance to the yellow hue, as defined in the
methods, ranged in the unit space. Based on this analyses, regions at
high risk of spillover are southeast Asia and
Madagascar.}\label{fig:risk}
}
\end{figure}

@@

Based on the previous result, we extracted the yellow component of the
risk map (TP add methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}.
However, this maps the potential risk, which must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et
al., 2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in
fig.~\ref{fig:compound}. Most of South America and Europe are at low
risk, as although densely populated, settlements tend to be in areas
with lower potential risk. However, this mapping reveals that South-East
Asia, the Indian subcontinent, and parts of sub-Saharan Africa, are at
high risk due to the overlap between built areas and bat communities
representing more opportunities for cross-species transmission of
betacoronaviruses.

\begin{figure}
\hypertarget{fig:compound}{%
\centering
\includegraphics{figures/risk_compounded.png}
\caption{Overlap of the percent of each pixel occupied by urbanized
structures, representing the degree of settlement, on the risk map.
Darker pixels correspond to more risk, in that the GMTC-derived risk of
fig.~\ref{fig:risk} is high \emph{and} the pixel is densely occupied by
human populations. This approach increases the relative risk of several
regions in Africa, and concenreates highlights the risk in India,
southeast China, and the arabic peninsula.}\label{fig:compound}
}
\end{figure}

@@

Our study focuses largely on the biogeography of hosts. Yet, we know
that viruses with high host plasticity, that is, the ability of a given
virus to adapt to various taxonomic orders and ecological groups
(Kreuder Johnson et al., 2015); are more likely to amplify viral
spillover, followed by secondary human-to-human transmission, and
geographical spread (Hazarie et al., 2021). High viral host plasticity
is an especially important trait for RNA viruses such as betacov
(Kreuder Johnson et al., 2015; Haddad et al., 2021). Indeed, our
analysis of viral sequences reveals that Latin America is a hotspot of
viral distinctiveness, suggesting that this part of the bats-betacov
system may be undergoing independent evolutionary dynamics (related
species sharing viruses that are different from the rest of the global
pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
richness is high but sharing is low; this suggests a different type of
evolutionary dynamics (unrelated viruses coevolving with evolutionarily
distinct hosts, generating high diversity locally). Both of these areas
should be priority areas for sampling, especially since
(\textbf{Becker?}) advance that they harbor undiscovered hosts of
beta-coronaviruses. This diversity of hosts, and the mechanisms by which
the exchange of viruses occurs between species, is largely affected by
the local environmental conditions and environmental change.

There are several factors that drive changes in the diversity of bats
(Alves et al., 2018), but human activities' effects on the ecosystem
(like modifications of land use) could significantly decrease it.
Therefore, it can be suggested that changes in the diversity of betacovs
in bats are linked to their biogeographic variation, and human
population density and other anthropogenic factors are decisive
moderators for its implications in public health. With the increase of
contact between humans and potential hosts, we also increase the risk of
emergence of novel diseases (Johnson et al., 2020), as previous studies
on RNA viruses suggest the importance of host phylogeography at the time
of virus dispersal (Gryseels et al., 2017). One of these scenarios where
interaction between bats and humans can occur can be seed dispersal in
tropical agroecosystems. It opens the discussion of whether the fruits
thrown by bats not only disperse seeds but could also be a source of
indirect interaction between viruses of bat origin and humans (Deshpande
et al., XX). This represents a challenge for conservation strategies and
disease ecology since we have areas with both potential for the
acquisition of zoonotic viruses and bat-human interaction; in
particular, the challenge lies in the fact that actual exposure must be
quantified from several scenarios, including both direct and indirect
bat - human interaction.

### 2022-04-30 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@038b5ced3967567216335ab65613b6abe255f450 🚀

Coming soon

>>>>

Driven by the need to understand the ecological factors involved in the
emergence of betacoronavirus through bat hosts, we develop an approach
to the assesment of spillover risk based on the Geographic Mosaic Theory
of Coevolution. In doing so, we provide a global mapping of the risk
posed by viruses, reflecting the fact that this risk has many ecological
and evolutionary origins. Our framework reveals that reservoir richness
alone is not a sufficient predictor of risk, and offers alternative
insights based on viral sharing, host compositional uniqueness, and host
phylogenetic diversity. By comparing our aggregated measure of risk to a
proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
system may originate novel viruses in humans.
